### 2016-01-23 AcademicPaperSummaries/PaperSummaries.tex
Added Chenoweth and Drakos & Gofas

\begin{document}
1/16/16 Initialization
this is a document to provide 1-2 paragraph writeups of papers that I have read in classes. Short-term goal is for quals prep.
the document gets backed up to https://github.com/margaretfoster/PaperSummaries, so that I can have Beeminder track my quals prep 
while I develop good habits.

@@

\underline{Research question and research area}
\underline{Summary of argument and mechanism}
\underline{Data \& methods}
\underline{Findings}
\underline{Notes on related papers}

@@

\underline{Research question and area:} McAdam and Paulsen connect to the literature of mobilization into high-risk political organizations. They primarily connect with the sociology literature, and, for example, don't make reference to free-rider problems that preoccupy the civil wars perspective on this question.
\underline{Summary of argument and mechanism:} McAdam and Paulsen are testing theories about how interpersonal ties and membership in organizations influence identity-based mobilization. The theory holds that social networks reinforce identity-based recruitment appeals by affirming that participation would support salient identity.
\underline{Data \& methods} Social network data generated from 1068 applications for participation in the 1964 Mississippi Freedom Summer project, coded networks based on surveys from participants, rejections, and no-shows and 556 followup surveys. Network-related independent variables operationalized as binary variables coding presence of listed organization or source of support.  Dependent variable was binary for participation in freedom summer; logit (?) statistical model.
\underline{Findings:} Personal ties are most important as sources of influence (vs conduits for information); suggestion that best networks for mobilization is weak ties between lots of tightly linked local organizations. Organizational support more important than individual ties for encouraging participation in a collective action.
\underline{Notes on related papers:} Lots of applicability here to mobilization for violent extremist organizations.

@@

Research question and research area: Blomberg et al. contribute to the literature on how economic conditions influence terrorist activities.
Terrorism definition: differentiate between ``rebellion'' which happens when a group that wants to change the status quo overthrows the government and takes power and ``terrorist attacks'' when dissident groups use ``terror activities'' to ``increase voice in economy'' but are not able to overthrow government. Note the almost tautological definition and the build-in assumption that terror attacks are intended to express a perspective on economy. Terror attacks conceptualized as ``low-intensity attacks'' designed to signal discontent with status quo. [This definition would include the far-left / far-right vandalism.]
\underline{Summary of argument and mechanism:} Groups choose between rebellion and terror attacks according to whether the country can give in to the dissidents. Strong institutions attracts terrorism; weak ones attract rebellion. [Observe that this outcome is built into their definition: weaker states should disproportionately attract attacks that threaten the state institutions because the institutions are more vulnerable.] They have a strong underlying assumption that attacks are intended to signal discontent. Assumption that attacks are either signals of anger or earnest attempts to overthrow government.
\underline{Data \& methods:} Dataset of 130 countries from 1968-1991, country-year unit of analysis. Data from ITERATE, which is transnational attacks and thus doesn't really match their theory, with economic indicators. Build a model of economic state change and terrorism/no terrorism through Markov processes. Do not appear to have any measure of the rebellion actions of their theory and look at conditional probabilities of state change. [Methodological question: are Markov chains at all appropriate for strategic interactions? In general, they are only applicable for linear systems.]
\underline{Findings:} In democratic, high income, countries, recessions increase probabilities of terror attacks. 
Notes on related papers:

@@

\underline{Research question and research area}
\underline{Summary of argument and mechanism}
\underline{Data \& methods}
\underline{Findings}
\underline{Notes on related papers}

>>>>

\begin{document}
1/16/16 Initialization
this is a document to provide 1-2 paragraph writeups of papers that I have read in classes. Short-term goal is for quals prep.
the document gets backed up to \\
https://github.com/margaretfoster/PaperSummaries, so that I can have Beeminder track my quals prep 
while I develop good habits.

@@

\underline{Research question and research area}\\
\underline{Summary of argument and mechanism}\\
\underline{Data \& methods}\\
\underline{Findings}\\
\underline{Notes on related papers}\\

@@

\underline{Research question and area:} McAdam and Paulsen connect to the literature of mobilization into high-risk political organizations. They primarily connect with the sociology literature, and, for example, don't make reference to free-rider problems that preoccupy the civil wars perspective on this question.\\
\underline{Summary of argument and mechanism:} McAdam and Paulsen are testing theories about how interpersonal ties and membership in organizations influence identity-based mobilization. The theory holds that social networks reinforce identity-based recruitment appeals by affirming that participation would support salient identity.
\underline{Data \& methods} Social network data generated from 1068 applications for participation in the 1964 Mississippi Freedom Summer project, coded networks based on surveys from participants, rejections, and no-shows and 556 followup surveys. Network-related independent variables operationalized as binary variables coding presence of listed organization or source of support.  Dependent variable was binary for participation in freedom summer; logit (?) statistical model.
\underline{Findings:} Personal ties are most important as sources of influence (vs conduits for information); suggestion that best networks for mobilization is weak ties between lots of tightly linked local organizations. Organizational support more important than individual ties for encouraging participation in a collective action.
\underline{Notes on related papers:} Lots of applicability here to mobilization for violent extremist organizations.

@@

\underline{Research question and area:} Blomberg et al. contribute to the literature on how economic conditions influence terrorist activities.\\
Terrorism definition: differentiate between ``rebellion'' which happens when a group that wants to change the status quo overthrows the government and takes power and ``terrorist attacks'' when dissident groups use ``terror activities'' to ``increase voice in economy'' but are not able to overthrow government. Note the almost tautological definition and the build-in assumption that terror attacks are intended to express a perspective on economy. Terror attacks conceptualized as ``low-intensity attacks'' designed to signal discontent with status quo. [This definition would include the far-left / far-right vandalism.]
\underline{Summary of argument and mechanism:} Groups choose between rebellion and terror attacks according to whether the country can give in to the dissidents. Strong institutions attracts terrorism; weak ones attract rebellion. [Observe that this outcome is built into their definition: weaker states should disproportionately attract attacks that threaten the state institutions because the institutions are more vulnerable.] They have a strong underlying assumption that attacks are intended to signal discontent. Assumption that attacks are either signals of anger or earnest attempts to overthrow government.
\underline{Data \& methods:} Dataset of 130 countries from 1968-1991, country-year unit of analysis. Data from ITERATE, which is transnational attacks and thus doesn't really match their theory, with economic indicators. Build a model of economic state change and terrorism/no terrorism through Markov processes. Do not appear to have any measure of the rebellion actions of their theory and look at conditional probabilities of state change. [Methodological question: are Markov chains at all appropriate for strategic interactions? In general, they are only applicable for linear systems.]
\underline{Findings:} In democratic, high income, countries, recessions increase probabilities of terror attacks. 
Notes on related papers:

@@

\underline{Keywords:} literature review.
\underline{Research question and research area:} Chenoweth provides an overview of research on the question f whether democracies disproportionately attract terrorist violence. \\
\underline{Operational definition of terrorism:} Here Chenoweth works from the basis that terrorsim is the deliberate use or threat of force against noncombatants by a non-state actor in pursuit of a political goals.
\underline{Summary of argument and mechanism:} Surveys underlying arguments about why democracies might, or might not, attract terrorism that are found in the articles that she synthesizes. 
\underline{Data \& methods:} Varies based on paper surveyed, but in general the studies all based on observed attacks and the GTD features heavily in both the motivating introduction and in the studies that she discusses. 
\underline{Findings:} She synthesizes existing studies to argue that terrorism is not a strategy of desperation, in particular that there is a non-linear relationship, with more violence in anocracies and partial democracies. As well, the patterns of the data change after 2001, with terrorism shifting from democracies to partial democracies and authoritarian regimes
\underline{Notes on related papers:} As a general note- this is probably a major source, and likely even a template, for any question about democracies and terrorism.

\textbf{Drakos, Kostas and Andreas Gofas. 2006. “The Devil You Know but are Afraid to Face:
Underreporting Bias and its Distorting Effects on the Study of Terrorism.” Journal of Conflict Resolution 50(5):714-735}

\underline{Keywords:} Terrorism; methodology

\underline{Research question and research area:} This is primarily a methodological paper- Drakos and Gofas are concerned with underreporting bias over terror attacks, leading to the situation in which the count of attacks that we see in news is lower than full sample of terrorist attacks and systematically underrepresents attacks in autocracies without a free press. This paper also speaks to the democracy and terrorism literature, which is where they primarily orient their literature review.
\underline{Construction of terrorism:} Terrorism is violence, or threat of violence calculated to create an atmosphere of fear and alarm. [Observe that this allows for states to perpetuate terrorism, which might be particularly systematically underreported.] 
\underline{Summary of argument and mechanism:} Argue that bias depends on degree of press freedom in country, so less freedom means underpresentation in aggregate databases like the GTD and findings that democracies attract more terrorism. Argue that cannot just put in a control for press freedom in part because it is a component of Polity.
\underline{Data \& methods:} Use concept of ``thinning'' in which count-data is generated as the true number of events weighted by some country-specific probability of inclusion. Assume that policy is ``fundamental determinant'' of underreporting bias because of inclusion of press freedom. Data sample is country-year unit for 153 countries from 1985-1998. Data from MIPT and Rand. Determine underreporting by looking at differences between two linear models (pg 732 of article).
\underline{Findings:} Find evidence that they claim is strongly indicative of underreporting bias in many countries, but they avoid making country-specific observations.
\underline{Notes on related papers:} Wonder if one could use Bayesian techniques to extend this approach, for example, by saying that the posterior would be underreporting $|$ observed attacks.
